---
title: "THINGS-Kids: developmental oddity task"
subtitle: "Mixed-effects analysis of the museum kiosk and Prolific adult samples"
date: today
date-format: "D MMMM YYYY"
format:
  html:
    theme: [cosmo, custom.scss]
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    number-sections: true
    number-depth: 2
    embed-resources: true
    fig-align: center
    fig-width: 8
    fig-height: 4.6
    code-fold: true
    code-summary: "Show code"
    code-tools: false
    df-print: paged
    tbl-cap-location: top
    link-external-newwindow: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
params:
  sessions: "data/sessions.csv"
  trials: "data/trials.csv"
  raw: "data/raw.json"
  manifest: "../public/manifest.json"
  aoa: "trial_aoa.csv"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8, fig.height = 4.6, dpi = 150)
library(dplyr); library(tidyr); library(ggplot2); library(knitr)
library(jsonlite); library(lme4); library(lmerTest); library(broom.mixed)
library(patchwork)

theme_set(theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(colour = "grey35", size = 10.5),
        strip.text = element_text(face = "bold")))

PAL    <- c(visual_agree = "#2E7D8F", visual_conflict = "#C1666B")
ARMPAL <- c(children = "#B5651D", adults = "#2E7D8F")
CHANCE <- 1/3
RT_FLOOR <- 250   # ms; below this nobody has viewed three photographs
PROLIFIC_ID <- "^[0-9a-f]{24}$"

`%||%` <- function(a, b) if (is.null(a)) b else a
pct <- function(x, d = 1) sprintf(paste0("%.", d, "f%%"), 100 * x)
fp  <- function(p) if (is.na(p)) "NA" else if (p < .001) "< .001" else sub("^0", "", sprintf("%.3f", p))

sessions <- read.csv(params$sessions, stringsAsFactors = FALSE)
trials   <- read.csv(params$trials,   stringsAsFactors = FALSE)
as_bool <- function(x) { if (is.logical(x)) return(x)
  v <- tolower(trimws(as.character(x))); o <- rep(NA, length(x))
  o[v %in% c("true","1")] <- TRUE; o[v %in% c("false","0")] <- FALSE; as.logical(o) }
for (v in c("complete","complete_inferred","qa_pass", grep("^qa_", names(sessions), value = TRUE)))
  if (v %in% names(sessions)) sessions[[v]] <- as_bool(sessions[[v]])
trials$correct <- as.numeric(as_bool(trials$correct))

# ---- session classification, from raw.json (carries consent_version) --------
docs <- fromJSON(params$raw, simplifyDataFrame = FALSE)
sess_meta <- bind_rows(lapply(docs, function(x) {
  s <- x$screen; if (!is.list(s)) s <- list()
  cs <- x$consent; if (!is.list(cs)) cs <- list()
  data.frame(participantID = x$participantID,
             ua = x$ua %||% NA_character_,
             consent_version = x$consent_version %||% NA_character_,
             age_label = (cs$age %||% NA_character_),
             w = s$w %||% NA_real_, h = s$h %||% NA_real_,
             n_trials = x$n_trials %||% 0,
             complete = isTRUE(x$complete), stringsAsFactors = FALSE) })) %>%
  mutate(
    # The kiosk is an iPad Pro in portrait running iPad Safari. Our own probes
    # share the kid_* id scheme but arrive from desktop/headless Chrome.
    kiosk_device = grepl("605\\.1\\.15", ua) & !grepl("Chrome", ua) &
                   !is.na(w) & w == 1024 & h == 1366,
    is_prolific = grepl(PROLIFIC_ID, participantID),
    age = suppressWarnings(as.numeric(sub("\\+$", "", age_label))))

sessions <- sessions %>% left_join(
  select(sess_meta, participantID, kiosk_device, consent_version, age_label, age), by = "participantID")

# Adults: unchanged definition (finished + passed QA).
adults <- sessions %>%
  mutate(finished = (!is.na(complete) & complete) |
                    (!is.na(complete_inferred) & complete_inferred) | !is.na(finishedAt)) %>%
  filter(grepl(PROLIFIC_ID, participantID), finished, !is.na(qa_pass) & qa_pass)

# Children: EVERY kiosk session, complete or not. A child who did 8 trials is a
# child who did 8 trials, not a missing data point.
kids_all <- sess_meta %>% filter(kiosk_device, age_label != "adult" | is.na(age_label))
kids_started <- kids_all %>% filter(n_trials > 0)

# meta$session_length describes the CURRENT design. These children ran whatever
# design was live when they sat down, so completion has to come from each
# session's own `complete` flag, and the nominal length from the data itself.
# Keying off the manifest silently reports "0 finished" the moment the design
# changes.
RUN_LEN <- {
  done <- kids_all$n_trials[kids_all$complete]
  if (length(done)) as.integer(names(sort(table(done), decreasing = TRUE))[1])
  else meta$session_length
}

# ---- stimulus covariates ----------------------------------------------------
mf <- fromJSON(params$manifest, simplifyDataFrame = FALSE)
rows <- list()
addr <- function(tr) rows[[length(rows)+1]] <<- data.frame(
  trial_id = tr$trial_id, m_condition = tr$condition %||% NA_character_,
  p_correct_spose = tr$p_correct_spose %||% NA_real_,
  odd_concept = tr$odd_concept %||% NA_character_,
  odd_category = tr$odd_category %||% NA_character_, stringsAsFactors = FALSE)
for (k in c("intro","core")) for (tr in mf[[k]] %||% list()) addr(tr)
for (k in c("blocks","adult_blocks")) for (b in mf[[k]] %||% list()) for (tr in b) addr(tr)
items <- distinct(bind_rows(rows), trial_id, .keep_all = TRUE)
# Age-of-acquisition norms for every concept, carried in from the vocab-gradient
# pipeline (rendering/_raw/vocab-gradient/data/item_metadata). aoa_target is the
# AoA of the odd-one-out; aoa_max is the latest-acquired of the three concepts.
items <- left_join(items, read.csv(params$aoa) %>%
                     select(trial_id, aoa_target, aoa_max, aoa_mean), by = "trial_id")
meta <- mf$meta

prep <- function(pids) trials %>% filter(participantID %in% pids) %>%
  left_join(items, by = "trial_id") %>%
  mutate(condition = coalesce(m_condition, condition),
         is_test = tier %in% c("familiar","warmup"))

kd <- prep(kids_started$participantID) %>%
  left_join(select(kids_started, participantID, age, age_label), by = "participantID")
ad <- prep(adults$participantID) %>% mutate(age = NA_real_)

# Trial-level engagement floor. This replaces excluding whole participants:
# a child's fast taps are dropped, their considered responses are kept.
kdf <- kd %>% filter(rt >= RT_FLOOR)
adf <- ad %>% filter(rt >= RT_FLOOR)

# participant-level deliberation, for use as a covariate (never as a filter)
delib <- kd %>% group_by(participantID) %>%
  summarise(median_rt = median(rt), frac_fast = mean(rt < 500), n_done = n(), .groups="drop")
kdf <- left_join(kdf, delib, by = "participantID") %>%
  left_join(select(kids_started, participantID, completed = complete), by = "participantID")
kd  <- left_join(kd,  delib, by = "participantID")

# The basic-level tier arrived on 2026-08-21 and is a different kind of item, so
# it is held out of every vocab-gradient analysis and reported on its own. `kds`
# is the SPoSE bank only; `kdb` is the new tier.
kdb <- kdf %>% filter(tier == "basic_level")
kds <- kdf %>% filter(tier != "basic_level")
adb <- adf %>% filter(tier == "basic_level")
adf <- adf %>% filter(tier != "basic_level")

wilson <- function(k,n,z=1.96){ if(is.na(n)||n==0) return(c(NA,NA)); p<-k/n; d<-1+z^2/n
  c(max(0,(p+z^2/(2*n))/d - z*sqrt(p*(1-p)/n+z^2/(4*n^2))/d),
    min(1,(p+z^2/(2*n))/d + z*sqrt(p*(1-p)/n+z^2/(4*n^2))/d)) }
wlo <- function(k,n) vapply(seq_along(k), function(i) wilson(k[i],n[i])[1], numeric(1))
whi <- function(k,n) vapply(seq_along(k), function(i) wilson(k[i],n[i])[2], numeric(1))
```

```{r main-models}
GL <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 3e5))
# Primary model. Crossed random intercepts absorb (a) how able each child is and
# (b) how hard each item is — the latter matters because short sessions are
# disproportionately made of the easy training/warmup trials that come first.
m_age <- glmer(correct ~ scale(age) + scale(aoa_target) + is_test +
                 (1|participantID) + (1|trial_id),
               data = filter(kds, !is.na(aoa_target)), family = binomial, control = GL)
b_age <- tidy(m_age, effects = "fixed") %>% filter(term == "scale(age)")

n_kid_started <- nrow(kids_started); n_kid_all <- nrow(kids_all)
n_kid_complete <- sum(kids_started$complete)
n_ad <- nrow(adults)
```

::: {.callout-important appearance="simple"}
## Summary

**All `r n_kid_started` children who answered at least one trial are in this analysis, not
just the `r n_kid_complete` who finished.** This doubles the child sample and reverses the
result: accuracy rises with age at
**β = `r sprintf('%+.2f', b_age$estimate)`, *p* = `r fp(b_age$p.value)`** (logit, per SD of
age), where the completers-only analysis could not detect it (*p* ≈ .09). The effect is *stronger*, not weaker, among the most
engaged children, so it is not an artefact of young children tapping through.
:::

---

# Why partial sessions are included

A museum kiosk produces mostly partial sessions. Discarding them is not a neutral choice:

```{r bias-table}
kids_started %>%
  mutate(grp = ifelse(complete, "completers", "partial")) %>%
  group_by(grp) %>%
  summarise(children = n(), median_age = median(age, na.rm = TRUE),
            trials_contributed = sum(n_trials), .groups = "drop") %>%
  kable(caption = "Completers vs. partial sessions.", digits = 2)
```

```{r bias-plot, fig.height = 3.6}
kids_started %>%
  mutate(grp = ifelse(complete, "completers", "partial")) %>%
  count(age, grp) %>%
  ggplot(aes(x = factor(age), y = n, fill = grp)) +
  geom_col(position = "stack", width = 0.72) +
  scale_fill_manual(values = c(completers = "#B5651D", partial = "grey70"), name = NULL) +
  scale_x_discrete(labels = function(x) ifelse(x == "10", "10+", x)) +
  labs(title = "Dropping partial sessions silently drops the youngest children",
       subtitle = "Ages 8–9 finish almost always; ages 2–3 almost never",
       x = "age (years)", y = "children") +
  theme(legend.position = "top")
```

Completion is age-graded, so a completers-only sample is a systematically **older** sample.
That biases the very gradient the study is trying to measure. Two further reasons to keep
the partials:

- Short sessions are not random trials. The session runs training (trials 1–4) → warmup
  (5–8) → test (9+), so a child who stopped early contributes mostly *easy* items. Comparing
  raw accuracy across session lengths is therefore confounded. An item random intercept
  absorbs this, which is why the models below are mixed rather than aggregate.
- Unbalanced data is what mixed models are for. A child with 8 trials is partially pooled
  toward the group mean and contributes proportionally less than a child with 51, instead of
  either counting equally or being thrown away.

# Sample

```{r accounting}
sessions %>%
  mutate(bucket = case_when(
    participantID %in% adults$participantID ~ "Prolific adults (usable)",
    kiosk_device & age_label == "adult" ~ "Kiosk — accompanying adult",
    kiosk_device & n_trials > 0 ~ "Kiosk children — answered >= 1 trial",
    kiosk_device ~ "Kiosk children — consented, never started",
    grepl(PROLIFIC_ID, participantID) ~ "Prolific — failed QA",
    study == "things_kids_adult_pilot" ~ "In-lab pilot (child bank)",
    study == "deploy_smoke_test" ~ "Deploy probe",
    TRUE ~ "Our own dev sessions")) %>%
  count(bucket, name = "sessions") %>% arrange(desc(sessions)) %>%
  kable(caption = paste0("All ", nrow(sessions), " documents in the collection."))
```

```{r survival, fig.height = 3.8}
surv <- data.frame(k = 0:RUN_LEN) %>%
  rowwise() %>% mutate(still = sum(kids_all$n_trials >= k)) %>% ungroup()
p_s <- ggplot(surv, aes(k, still)) +
  geom_step(colour = "#B5651D", linewidth = 1) +
  geom_vline(xintercept = 8.5, linetype = "dashed", colour = "grey40") +
  annotate("text", x = 9.5, y = max(surv$still)*0.85, hjust = 0, size = 3.2,
           colour = "grey30", label = "training + warmup end") +
  labs(title = "Where children stop", x = "trials completed", y = "children still going",
       subtitle = paste0(n_kid_all, " children consented on the kiosk"))

comp_age <- kids_all %>% group_by(age) %>%
  summarise(n = n(), finished = sum(complete), .groups="drop") %>%
  mutate(rate = finished/n)
p_c <- ggplot(comp_age, aes(age, rate)) +
  geom_point(aes(size = n), colour = "#B5651D") +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),
              colour = "#B5651D", fill = "#B5651D", alpha = 0.15, se = TRUE) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = 2:10, labels = c(2:9, "10+")) +
  scale_size_continuous(range = c(2, 6), name = "children") +
  labs(title = "Completion rises with age", x = "age (years)", y = "% finishing the session")
p_s + p_c
```

```{r dropout-model}
md <- glm(complete ~ age, data = kids_all, family = binomial)
kable(broom::tidy(md), digits = 3,
      caption = "Logistic regression: finishing the full session ~ age.")
```

`r sum(kids_all$n_trials == 0)` children consented but never answered a trial. Among those who
started, the commonest stopping point is **trial 8**, the boundary where training and warmup
end and the test trials begin. Attrition is therefore concentrated at the transition to the
harder items rather than distributed across the session.

# Accuracy and age

Accuracy is binary, so these are mixed-effects **logistic** regressions (GLMMs): a linear
mixed model on the log-odds scale. Both grouping factors are crossed random intercepts,
child and item. Random slopes for condition by child were fitted first and were singular, so
the models below retain random intercepts only.

```{r age-model}
tidy(m_age, effects = "fixed", conf.int = TRUE) %>%
  transmute(term, estimate, std.error, odds_ratio = exp(estimate),
            conf.low = exp(conf.low), conf.high = exp(conf.high), p.value) %>%
  kable(digits = 3, caption = paste("correct ~ age + target AoA + tier + (1|child) + (1|item).",
        "Predictors are standardised, so estimates are per SD."))

vc <- as.data.frame(VarCorr(m_age))
kable(vc[, c("grp","sdcor")], digits = 3, col.names = c("random effect","SD (logit)"),
      caption = "Variance components: how much children differ, and how much items differ.")
```

```{r age-plot, fig.height = 4.6}
per_child <- kds %>% group_by(participantID, age, n_done) %>%
  summarise(n = n(), k = sum(correct), acc = mean(correct), .groups="drop")

p1 <- ggplot(per_child, aes(age, acc)) +
  geom_hline(yintercept = CHANCE, linetype = "dashed", colour = "grey55") +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),
              aes(weight = n), colour = "#B5651D", fill = "#B5651D", alpha = 0.15) +
  geom_point(aes(size = n), colour = "#B5651D", alpha = 0.75) +
  scale_size_continuous(range = c(1.5, 6), name = "trials") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = 2:10, labels = c(2:9,"10+")) +
  annotate("text", x = 2, y = CHANCE - 0.05, label = "chance", colour = "grey45",
           size = 3.2, hjust = 0) +
  labs(title = "Every child who answered a trial", x = "age (years)",
       y = "accuracy (trials >= 250 ms)",
       subtitle = "Point size = trials contributed; fit is trial-weighted")

band <- per_child %>% mutate(b = cut(age, c(1,4,6,11), labels = c("2–4","5–6","7–10+"))) %>%
  group_by(b) %>% summarise(children = n(), n = sum(n), k = sum(k), acc = sum(k)/sum(n), .groups="drop") %>%
  mutate(lo = wlo(k,n), hi = whi(k,n))
ad_acc <- mean(adf$correct[adf$is_test])
p2 <- ggplot(band, aes(b, acc)) +
  geom_hline(yintercept = CHANCE, linetype = "dashed", colour = "grey55") +
  geom_hline(yintercept = ad_acc, linetype = "dotted", colour = "#2E7D8F", linewidth = 0.8) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.12, colour = "grey40") +
  geom_point(size = 3.6, colour = "#B5651D") +
  annotate("text", x = 0.62, y = ad_acc + 0.035, label = "adults", colour = "#2E7D8F",
           size = 3.2, hjust = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0.2, 1)) +
  labs(title = "By age band", x = NULL, y = NULL, subtitle = "95% Wilson CI, pooled trials")
p1 + p2
```

```{r band-tab}
band %>% transmute(band = b, children, trials = n, accuracy = acc,
                   ci = sprintf("[%.3f, %.3f]", lo, hi)) %>%
  kable(digits = 3, caption = "Pooled accuracy by age band (all children, all trials).")
```

The child random-effect SD (`r round(vc$sdcor[vc$grp=="participantID"], 2)` logit) is larger
than the item SD (`r round(vc$sdcor[vc$grp=="trial_id"], 2)`): children differ from one
another more than items differ from one another. That is the variation the model is
accounting for, and the reason a fixed-effects comparison of group means would overstate its
own precision.

# Robustness to inclusion criteria

```{r robust}
spec <- function(lab, d) {
  d <- d %>% filter(!is.na(age), !is.na(aoa_target))
  if (n_distinct(d$participantID) < 6) return(NULL)
  m <- glmer(correct ~ scale(age) + scale(aoa_target) + is_test + (1|participantID) + (1|trial_id),
             data = filter(d, !is.na(aoa_target)), family = binomial, control = GL)
  r <- tidy(m, effects = "fixed") %>% filter(term == "scale(age)")
  data.frame(specification = lab, children = n_distinct(d$participantID), trials = nrow(d),
             beta = r$estimate, se = r$std.error, p = r$p.value)
}
rob <- bind_rows(
  spec("ALL trials, no RT floor",              filter(kd, tier != "basic_level")),
  spec("RT floor 250 ms (primary)",            kds),
  spec("RT floor 500 ms",                      filter(kd, rt >= 500, tier != "basic_level")),
  spec("most-deliberate children only",        filter(kds, median_rt >= 800)),
  spec("completers only",                      filter(kds, completed)),
  spec("completers + deliberate (old report)", filter(kds, completed, median_rt >= 800)))
rob %>% mutate(p = vapply(p, fp, character(1))) %>%
  kable(digits = 3, caption = "The age effect under six inclusion rules.")
```

```{r robust-plot, fig.height = 3.4}
ggplot(rob, aes(x = reorder(specification, beta), y = beta)) +
  geom_hline(yintercept = 0, colour = "grey45") +
  geom_errorbar(aes(ymin = beta - 1.96*se, ymax = beta + 1.96*se), width = 0.14, colour = "grey40") +
  geom_point(aes(colour = p < .05), size = 3.2) +
  scale_colour_manual(values = c(`TRUE` = "#B5651D", `FALSE` = "grey65"),
                      labels = c(`TRUE` = "p < .05", `FALSE` = "n.s."), name = NULL) +
  coord_flip() +
  labs(title = "The age effect is a power problem, not a robustness problem",
       subtitle = "Point estimate is stable across inclusion rules; only the CI changes",
       x = NULL, y = "β for age (logit, per SD)")
```

The estimate is stable across specifications, every one landing near
β = `r sprintf('%.2f', mean(rob$beta))`. What changes is the width of the interval. The two
completers-only rows correspond to the earlier analysis; they are non-significant because
they discard half the children, not because the effect differs.

```{r tap-check}
tapper <- kids_started %>% left_join(delib, by = "participantID") %>%
  mutate(taps = median_rt < 800)
ct <- suppressWarnings(cor.test(tapper$age, as.numeric(tapper$taps)))
m_cov <- glmer(correct ~ scale(age) + scale(log(median_rt)) + is_test +
                 (1|participantID) + (1|trial_id), data = kds, family = binomial, control = GL)
tidy(m_cov, effects = "fixed") %>% transmute(term, estimate, std.error, p.value) %>%
  kable(digits = 3, caption = paste("Age effect controlling for how deliberate each child was.",
        "Both predictors are independently reliable."))
```

The obvious worry is circular: if young children tap through, low accuracy would follow from
age mechanically. It does not hold. Tapping is unrelated to age
(*r* = `r round(ct$estimate, 2)`, *p* = `r fp(ct$p.value)`; mean age
`r round(mean(tapper$age[tapper$taps], na.rm=TRUE), 1)` for tappers vs
`r round(mean(tapper$age[!tapper$taps], na.rm=TRUE), 1)`), the age effect *strengthens* when
only deliberate children are kept, and age survives controlling for deliberation directly.

# Vocabulary as an alternative account

An oddity trial is unanswerable if the child does not know the target concept, so an age
effect could be nothing more than older children knowing more words. Every concept carries an
estimated age-of-acquisition norm (in years), which lets that be tested directly rather than
assumed away. Target AoA spans `r sprintf('%.1f', min(items$aoa_target, na.rm=TRUE))`–`r sprintf('%.1f', max(items$aoa_target, na.rm=TRUE))`
years, so there is real range to work with.

```{r aoa-models}
kv <- kds %>% filter(!is.na(aoa_target))
mk <- function(f, d = kv) glmer(as.formula(f), data = d, family = binomial, control = GL)
grab <- function(lab, m, terms) {
  tidy(m, effects = "fixed") %>% filter(term %in% terms) %>%
    transmute(model = lab, term, estimate, std.error, p.value) }
aoa_tab <- bind_rows(
  grab("A. age alone", mk("correct~scale(age)+is_test+(1|participantID)+(1|trial_id)"), "scale(age)"),
  grab("B. AoA alone", mk("correct~scale(aoa_target)+is_test+(1|participantID)+(1|trial_id)"), "scale(aoa_target)"),
  grab("C. age + AoA", mk("correct~scale(age)+scale(aoa_target)+is_test+(1|participantID)+(1|trial_id)"),
       c("scale(age)","scale(aoa_target)")),
  grab("D. age x AoA", mk("correct~scale(age)*scale(aoa_target)+is_test+(1|participantID)+(1|trial_id)"),
       "scale(age):scale(aoa_target)"),
  grab("E. + SPoSE difficulty", mk("correct~scale(age)+scale(aoa_target)+scale(p_correct_spose)+is_test+(1|participantID)+(1|trial_id)",
       filter(kv, !is.na(p_correct_spose))), c("scale(age)","scale(aoa_target)","scale(p_correct_spose)")),
  grab("F. adults, AoA", mk("correct~scale(aoa_target)+is_test+(1|participantID)+(1|trial_id)",
       filter(adf, !is.na(aoa_target))), "scale(aoa_target)"))
aoa_tab %>% mutate(p.value = vapply(p.value, fp, character(1))) %>%
  kable(digits = 3, caption = "Age and target AoA as competing accounts of child accuracy.")
```

```{r aoa-plot, fig.height = 4.2}
q_aoa <- kv %>% mutate(q = ntile(aoa_target, 4)) %>%
  group_by(q) %>% summarise(n = n(), aoa = mean(aoa_target), k = sum(correct),
                            acc = mean(correct), .groups="drop") %>%
  mutate(lo = wlo(k,n), hi = whi(k,n))
pa <- ggplot(q_aoa, aes(aoa, acc)) +
  geom_hline(yintercept = CHANCE, linetype = "dashed", colour = "grey55") +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.18, colour = "grey40") +
  geom_line(colour = "#B5651D", linewidth = 0.9) +
  geom_point(size = 3.4, colour = "#B5651D") +
  scale_y_continuous(labels = scales::percent, limits = c(0.3, 1)) +
  labs(title = "Accuracy is flat across target AoA", x = "mean target AoA of quartile (years)",
       y = "child accuracy", subtitle = "If vocabulary drove performance this would slope down")

band_aoa <- kv %>% mutate(q = ntile(aoa_target, 4),
                          band = cut(age, c(1,4,6,11), labels = c("2–4","5–6","7–10+"))) %>%
  group_by(band, q) %>% summarise(aoa = mean(aoa_target), acc = mean(correct), n = n(), .groups="drop")
pb <- ggplot(band_aoa, aes(aoa, acc, colour = band)) +
  geom_hline(yintercept = CHANCE, linetype = "dashed", colour = "grey55") +
  geom_line(linewidth = 0.9) + geom_point(size = 2.8) +
  scale_colour_manual(values = c("2–4"="#E0A458","5–6"="#B5651D","7–10+"="#6B3410"), name = "age") +
  scale_y_continuous(labels = scales::percent, limits = c(0.3, 1)) +
  labs(title = "And flat within every age band", x = "mean target AoA of quartile (years)",
       y = NULL, subtitle = "The bands are separated by age, not by AoA") +
  theme(legend.position = "top")
pa + pb
```

```{r aoa-numbers}
a_c <- aoa_tab %>% filter(model == "C. age + AoA")
a_age_c <- a_c %>% filter(term == "scale(age)")
a_aoa_c <- a_c %>% filter(term == "scale(aoa_target)")
a_int <- aoa_tab %>% filter(model == "D. age x AoA")
a_alone <- aoa_tab %>% filter(model == "A. age alone")
```

**No, it is not vocabulary.** Target AoA does not predict child accuracy on its own
(β = `r sprintf('%+.3f', aoa_tab$estimate[aoa_tab$model=="B. AoA alone"])`,
*p* = `r fp(aoa_tab$p.value[aoa_tab$model=="B. AoA alone"])`), and adding it leaves the age
effect unchanged: β = `r sprintf('%.3f', a_alone$estimate)` alone versus
`r sprintf('%.3f', a_age_c$estimate)` with AoA in the model. The two variables are not
competing for the same variance.

The interaction is the sharper test: if late-acquired words were the obstacle, they should
penalise 3-year-olds far more than 9-year-olds. The age × AoA term is
β = `r sprintf('%+.3f', a_int$estimate)` (*p* = `r fp(a_int$p.value)`), indistinguishable
from zero, and the age bands run parallel rather than fanning out. Adults show no AoA effect
either,
(*p* = `r fp(aoa_tab$p.value[aoa_tab$model=="F. adults, AoA"])`).

Meanwhile SPoSE difficulty *does* survive alongside both
(*p* = `r fp(aoa_tab$p.value[aoa_tab$model=="E. + SPoSE difficulty"] [aoa_tab$term[aoa_tab$model=="E. + SPoSE difficulty"]=="scale(p_correct_spose)"])`),
so item difficulty is measurable in these data. It is specifically the *word-learning*
account that fails.

Two limits apply. The task is picture-based and never asks a child to name anything, so
word-AoA is an imperfect proxy for whether a child holds the concept; and these are
adult-rated norms. Within this design, however, the vocabulary explanation has been given a fair
test and does not account for the age effect, which locates the developmental change in the
conceptual comparison rather than at
knowing the words.

## Which AoA should gate a trial

`aoa_target` is the odd-one-out alone, but an oddity judgement arguably needs all three
concepts: the child must see that the pair goes together *and* that the target does not. So
the binding constraint may be the **latest-acquired** of the three (`aoa_max`, mean
`r sprintf('%.1f', mean(items$aoa_max, na.rm=TRUE))` years) rather than the target's own AoA.
The two are only moderately related (*r* = `r sprintf('%.2f', cor(kv$aoa_target, kv$aoa_max))`),
so this is an independent test rather than a restatement.

```{r aoa-sens}
sens_aoa <- bind_rows(
  grab("target AoA", mk("correct~scale(age)+scale(aoa_target)+is_test+(1|participantID)+(1|trial_id)"),
       c("scale(age)","scale(aoa_target)")),
  grab("max AoA of triplet", mk("correct~scale(age)+scale(aoa_max)+is_test+(1|participantID)+(1|trial_id)"),
       c("scale(age)","scale(aoa_max)")),
  grab("mean AoA of triplet", mk("correct~scale(age)+scale(aoa_mean)+is_test+(1|participantID)+(1|trial_id)"),
       c("scale(age)","scale(aoa_mean)")),
  grab("max AoA x age", mk("correct~scale(age)*scale(aoa_max)+is_test+(1|participantID)+(1|trial_id)"),
       "scale(age):scale(aoa_max)"),
  grab("mean AoA x age", mk("correct~scale(age)*scale(aoa_mean)+is_test+(1|participantID)+(1|trial_id)"),
       "scale(age):scale(aoa_mean)"))
sens_aoa %>% mutate(p.value = vapply(p.value, fp, character(1))) %>%
  kable(digits = 3, caption = paste("The same null under three ways of scoring a trial's AoA.",
        "The age estimate is unchanged in every row."))
```

```{r aoa-sens-plot, fig.height = 3.8}
qmax <- kv %>% mutate(q = ntile(aoa_max, 4),
                      band = cut(age, c(1,4,6,11), labels = c("2–4","5–6","7–10+"))) %>%
  group_by(band, q) %>% summarise(aoa = mean(aoa_max), acc = mean(correct), n = n(), .groups="drop")
ggplot(qmax, aes(aoa, acc, colour = band)) +
  geom_hline(yintercept = CHANCE, linetype = "dashed", colour = "grey55") +
  geom_line(linewidth = 0.9) + geom_point(size = 2.8) +
  scale_colour_manual(values = c("2–4"="#E0A458","5–6"="#B5651D","7–10+"="#6B3410"), name = "age") +
  scale_y_continuous(labels = scales::percent, limits = c(0.3, 1)) +
  labs(title = "Even the hardest-vocabulary trials show no age-specific penalty",
       subtitle = "Lines stay parallel: late-acquired triplets do not punish the youngest children",
       x = "mean latest-acquired AoA of quartile (years)", y = "accuracy")
```

```{r aoa-sens-numbers}
smax <- sens_aoa %>% filter(model == "max AoA of triplet", term == "scale(aoa_max)")
sint <- sens_aoa %>% filter(model == "max AoA x age")
top_q <- qmax %>% filter(q == max(q))
```

The null survives intact. Scoring each trial by its latest-acquired concept gives
β = `r sprintf('%+.3f', smax$estimate)` (*p* = `r fp(smax$p.value)`), the interaction with age
is *p* = `r fp(sint$p.value)`, and the age coefficient stays at
`r sprintf('%.3f', sens_aoa$estimate[sens_aoa$term=="scale(age)"][1])` in all three
specifications. The top quartile averages `r sprintf('%.1f', mean(top_q$aoa))`-year AoA,
later than almost every child tested, yet 2–4 year-olds answer it at
`r pct(top_q$acc[top_q$band=="2–4"])`, their *best* quartile rather than their worst.

That is about as clean as a null gets: three operationalisations, no main effect, no
interaction, and no movement in the age estimate.

# The developmental prediction

```{r cond-model}
cond_all <- bind_rows(
  kds %>% filter(condition %in% c("visual_agree","visual_conflict")) %>% mutate(grp = "children"),
  adf %>% filter(condition %in% c("visual_agree","visual_conflict")) %>% mutate(grp = "adults")) %>%
  mutate(grp = factor(grp, levels = c("adults","children")))

m_cond <- glmer(correct ~ condition * grp + (1|participantID) + (1|trial_id),
                data = cond_all, family = binomial, control = GL)
tidy(m_cond, effects = "fixed", conf.int = TRUE) %>%
  transmute(term, estimate, std.error, odds_ratio = exp(estimate), p.value) %>%
  kable(digits = 3, caption = "correct ~ condition × group, crossed random effects.")

m_ca <- glmer(correct ~ condition * scale(age) + (1|participantID) + (1|trial_id),
              data = filter(cond_all, grp == "children"), family = binomial, control = GL)
tidy(m_ca, effects = "fixed") %>% transmute(term, estimate, std.error, p.value) %>%
  kable(digits = 3, caption = "Within children: does the conflict cost shrink with age?")
```

```{r cond-plot, fig.height = 4.4}
cs <- cond_all %>% group_by(grp, condition) %>%
  summarise(n = n(), k = sum(correct), acc = mean(correct), .groups="drop") %>%
  mutate(lo = wlo(k,n), hi = whi(k,n))
p3 <- ggplot(cs, aes(grp, acc, colour = condition, group = condition)) +
  geom_hline(yintercept = CHANCE, linetype = "dashed", colour = "grey55") +
  geom_line(linewidth = 0.9) +
  geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.08, linewidth = 0.7) +
  geom_point(size = 3.4) +
  scale_colour_manual(values = PAL, name = NULL,
                      labels = c("appearance agrees","appearance conflicts")) +
  scale_y_continuous(labels = scales::percent, limits = c(0.3, 1)) +
  labs(title = "Both conditions, both groups", x = NULL, y = "accuracy") +
  theme(legend.position = "top")

cost_age <- kds %>% filter(condition %in% c("visual_agree","visual_conflict")) %>%
  group_by(participantID, age, condition) %>% summarise(a = mean(correct), n = n(), .groups="drop") %>%
  group_by(participantID, age) %>% filter(n() == 2, min(n) >= 4) %>%
  summarise(cost = a[condition=="visual_agree"] - a[condition=="visual_conflict"],
            n = sum(n), .groups="drop")
p4 <- ggplot(cost_age, aes(age, cost)) +
  geom_hline(yintercept = 0, colour = "grey45") +
  geom_smooth(method = "lm", aes(weight = n), colour = "#B5651D", fill = "#B5651D", alpha = 0.15) +
  geom_point(aes(size = n), colour = "#B5651D", alpha = 0.75) +
  scale_size_continuous(range = c(1.5,5), name = "trials") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = 2:10, labels = c(2:9,"10+")) +
  labs(title = "Conflict cost by age", x = "age (years)",
       y = "accuracy(agree) − accuracy(conflict)",
       subtitle = paste(nrow(cost_age), "children with >= 4 trials in each condition"))
p3 + p4
```

```{r cond-numbers}
i_p <- tidy(m_cond, effects="fixed") %>% filter(grepl(":", term)) %>% pull(p.value)
ca_p <- tidy(m_ca, effects="fixed") %>% filter(grepl(":", term)) %>% pull(p.value)
cc <- cs %>% group_by(grp) %>% summarise(cost = acc[condition=="visual_agree"] - acc[condition=="visual_conflict"])
```

```{r floor-sens}
acc_by <- kds %>% group_by(participantID) %>% summarise(a = mean(correct), .groups="drop")
sens <- function(lab, keep) {
  kk <- kds %>% filter(participantID %in% keep, condition %in% c("visual_agree","visual_conflict"))
  d <- bind_rows(kk %>% mutate(grp="children"),
                 adf %>% filter(condition %in% c("visual_agree","visual_conflict")) %>% mutate(grp="adults")) %>%
       mutate(grp = factor(grp, levels=c("adults","children")))
  mm <- glmer(correct ~ condition*grp + (1|participantID) + (1|trial_id),
              data = d, family = binomial, control = GL)
  r <- tidy(mm, effects="fixed") %>% filter(grepl(":", term))
  data.frame(subset = lab, children = n_distinct(kk$participantID),
             child_cost = mean(kk$correct[kk$condition=="visual_agree"]) -
                          mean(kk$correct[kk$condition=="visual_conflict"]),
             interaction_beta = r$estimate, p = r$p.value)
}
fs <- bind_rows(
  sens("all children",                    acc_by$participantID),
  sens("accuracy > .40 (above chance)",   acc_by$participantID[acc_by$a > .40]),
  sens("accuracy > .50",                  acc_by$participantID[acc_by$a > .50]),
  sens("accuracy > .60 (adult-like)",     acc_by$participantID[acc_by$a > .60]))
fs %>% mutate(p = vapply(p, fp, character(1))) %>%
  kable(digits = 3, caption = paste("Floor sensitivity. A child at chance in both conditions",
        "has zero cost by construction, so the near-chance children are removed stepwise."))
```

Children lose `r pct(cc$cost[cc$grp=="children"])` to visual conflict, adults
`r pct(cc$cost[cc$grp=="adults"])`, and the group × condition interaction is
*p* = `r fp(i_p)`. **This is significant in the direction opposite to the prediction:
children pay a *smaller* appearance penalty than adults, not a larger one.**

That result needs one caveat taken seriously. Chance on this task is 1/3, not 0, and a child
responding at chance in both conditions has a zero cost by construction, so including
near-chance children mechanically shrinks the child estimate. The table above removes them
stepwise. The child cost does grow as they are dropped
(`r pct(fs$child_cost[1])` → `r pct(fs$child_cost[nrow(fs)])`), but the interaction estimate
barely moves (β `r sprintf('%.2f', fs$interaction_beta[1])` →
`r sprintf('%.2f', fs$interaction_beta[nrow(fs)])`); it loses significance because the sample
shrinks, not because the effect does. Even among children scoring in the adult range the cost
is `r pct(fs$child_cost[nrow(fs)])` against the adults' `r pct(cc$cost[cc$grp=="adults"])`.

The reading supported by these data is **a stable reversal that is not yet established**: the point estimate
consistently says children weight appearance *less* than adults, compression explains part but
not all of it, and within children the cost does not change with age (*p* = `r fp(ca_p)`).
If it holds, the study's founding prediction is inverted, which warrants
adequate powering rather than settling at *n* = `r n_kid_started`.

# Response times

RT is continuous, so this is an ordinary LMM on log RT, with the same crossed random effects.

```{r rt-model}
rt_d <- kds %>% filter(rt < 30000, condition %in% c("visual_agree","visual_conflict"))
m_rt <- lmer(log(rt) ~ condition + scale(age) + correct + (1|participantID) + (1|trial_id), data = rt_d)
tidy(m_rt, effects = "fixed") %>% transmute(term, estimate, std.error, df, statistic, p.value) %>%
  kable(digits = 3, caption = "log(RT) ~ condition + age + accuracy, children.")
```

```{r rt-plot, fig.height = 4}
p5 <- ggplot(rt_d, aes(rt/1000, fill = condition)) +
  geom_density(alpha = 0.45, colour = NA) + scale_x_log10() +
  scale_fill_manual(values = PAL, name = NULL,
                    labels = c("appearance agrees","appearance conflicts")) +
  labs(title = "Child RT distribution", x = "response time (s, log)", y = "density",
       subtitle = paste0("trials >= ", RT_FLOOR, " ms")) +
  theme(legend.position = "top")
p6 <- kds %>% group_by(participantID, age) %>% summarise(mrt = median(rt), .groups="drop") %>%
  ggplot(aes(age, mrt/1000)) +
  geom_point(colour = "#B5651D", size = 2.6, alpha = 0.8) +
  geom_smooth(method = "lm", colour = "#B5651D", fill = "#B5651D", alpha = 0.15) +
  scale_x_continuous(breaks = 2:10, labels = c(2:9,"10+")) +
  labs(title = "Median RT by age", x = "age (years)", y = "median RT (s)",
       subtitle = "Older children are not simply slower or faster")
p5 + p6
```

# SPoSE difficulty in children

```{r spose}
sp <- kds %>% filter(!is.na(p_correct_spose))
m_sp <- glmer(correct ~ scale(p_correct_spose) + scale(age) + (1|participantID) + (1|trial_id),
              data = sp, family = binomial, control = GL)
tidy(m_sp, effects="fixed") %>% transmute(term, estimate, std.error, odds_ratio = exp(estimate), p.value) %>%
  kable(digits = 3, caption = "Children: correct ~ SPoSE prediction + age.")
```

```{r spose-plot, fig.height = 3.8}
byq <- bind_rows(sp %>% mutate(grp="children"),
                 adf %>% filter(!is.na(p_correct_spose)) %>% mutate(grp="adults")) %>%
  group_by(grp) %>% mutate(q = ntile(p_correct_spose, 4)) %>%
  group_by(grp, q) %>% summarise(n=n(), predicted=mean(p_correct_spose), k=sum(correct),
                                 observed=mean(correct), .groups="drop") %>%
  mutate(lo=wlo(k,n), hi=whi(k,n))
ggplot(byq, aes(predicted, observed, colour = grp)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", colour="grey60") +
  geom_errorbar(aes(ymin=lo, ymax=hi), width=0.012, alpha=0.7) +
  geom_line(linewidth=0.8) + geom_point(size=3.2) +
  scale_colour_manual(values=c(adults="#2E7D8F", children="#B5651D"), name=NULL) +
  scale_x_continuous(labels=scales::percent) + scale_y_continuous(labels=scales::percent) +
  labs(title="SPoSE ranks difficulty for both groups, but is not calibrated",
       subtitle="Dashed line is perfect calibration",
       x="SPoSE-predicted accuracy", y="observed accuracy")
```

```{r hardest}
sp %>% group_by(trial_id) %>%
  summarise(n=n(), observed=mean(correct), predicted=first(p_correct_spose),
            odd=first(odd_concept), category=first(odd_category), .groups="drop") %>%
  filter(n >= 4) %>% arrange(observed) %>% head(12) %>%
  kable(digits=3, caption="Hardest child items (>= 4 observations, all children).")
```

## Item-level predictions and observed performance

The quartile view above averages over the model's errors. This section reports the
item-level version, with every well-observed trial as its own point, split by age.

```{r item-density}
item_age <- kds %>% filter(!is.na(p_correct_spose), !is.na(age)) %>%
  mutate(band = cut(age, c(1,4,6,11), labels = c("2–4","5–6","7–10+")))
dens <- item_age %>% count(trial_id, band) %>%
  tidyr::pivot_wider(names_from = band, values_from = n, values_fill = 0)
MIN_OBS <- 5
well <- dens %>% filter(dplyr::if_all(-trial_id, ~ .x >= MIN_OBS)) %>% pull(trial_id)
data.frame(
  quantity = c("trials children have seen", "median observations per trial",
               paste0("trials with >= ", MIN_OBS, " observations in EVERY age band"),
               "what those trials are"),
  value = c(n_distinct(item_age$trial_id),
            median(count(item_age, trial_id)$n), length(well),
            "the core + warm-up set, which every child receives")
) %>% kable(caption = "Why this section uses a subset.")
```

A rotating bank buys breadth, not depth: the median trial has been seen
`r median(count(item_age, trial_id)$n)` times, which cannot support an item-level estimate,
let alone one split three ways by age. Only the **`r length(well)` core and warm-up trials**,
the ones every child receives, carry enough observations. This section uses those.

```{r item-scatter, fig.height = 4.6}
ia <- item_age %>% filter(trial_id %in% well) %>%
  group_by(trial_id, band) %>%
  summarise(n = n(), k = sum(correct), observed = mean(correct),
            predicted = first(p_correct_spose), odd = first(odd_concept), .groups = "drop") %>%
  mutate(lo = wlo(k, n), hi = whi(k, n))

ggplot(ia, aes(predicted, observed)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey60") +
  geom_hline(yintercept = CHANCE, linetype = "dotted", colour = "grey60") +
  geom_linerange(aes(ymin = lo, ymax = hi), colour = "grey75", linewidth = 0.5) +
  geom_point(aes(size = n), colour = "#B5651D", alpha = 0.85) +
  geom_smooth(method = "lm", se = FALSE, colour = "#6B3410", linewidth = 0.8) +
  facet_wrap(~band) +
  scale_size_continuous(range = c(1.5, 4.5), name = "obs") +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(title = "Each point is one trial, at one age band",
       subtitle = paste("Dashed = perfect calibration, dotted = chance.",
                        "Points above the dashed line are trials children find",
                        "easier than SPoSE predicts."),
       x = "SPoSE-predicted accuracy", y = "observed child accuracy")
```

```{r item-corr}
ia %>% group_by(band) %>%
  summarise(trials = n(), r = cor(predicted, observed),
            mean_predicted = mean(predicted), mean_observed = mean(observed),
            .groups = "drop") %>%
  mutate(calibration_gap = mean_observed - mean_predicted) %>%
  kable(digits = 3, caption = paste("Item-level correlation between prediction and",
        "performance, within each age band."))
```

```{r item-slope, fig.height = 4.4}
wide <- ia %>% select(trial_id, odd, band, observed, predicted) %>%
  tidyr::pivot_wider(names_from = band, values_from = observed)
lab <- wide %>% mutate(spread = `7–10+` - `2–4`) %>%
  arrange(spread) %>% slice(c(1:3, (n()-2):n()))

ggplot(ia, aes(band, observed, group = trial_id)) +
  geom_hline(yintercept = CHANCE, linetype = "dotted", colour = "grey60") +
  geom_line(colour = "grey72", linewidth = 0.5) +
  geom_point(aes(colour = predicted), size = 2.4) +
  scale_colour_gradient(low = "#C1666B", high = "#2E7D8F",
                        labels = scales::percent, name = "SPoSE\npredicted") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(title = "The same trial across ages", x = NULL, y = "observed accuracy",
       subtitle = "One line per trial; colour is what SPoSE predicted for it")
```

```{r item-table}
wide %>%
  transmute(trial_id, odd_concept = odd, SPoSE = predicted,
            `2–4` = `2–4`, `5–6` = `5–6`, `7–10+` = `7–10+`,
            gap_young = `2–4` - predicted) %>%
  arrange(gap_young) %>%
  kable(digits = 3, caption = paste("Every well-observed trial: what SPoSE predicted and",
        "what each age band actually did. Sorted by how badly the model over-predicts",
        "the youngest children."))
```

```{r item-numbers}
cr <- ia %>% group_by(band) %>% summarise(r = cor(predicted, observed),
        gap = mean(observed) - mean(predicted), .groups="drop")
ranked <- wide %>% mutate(g = `2–4` - predicted) %>% arrange(g)
worst <- ranked %>% slice(1); best <- ranked %>% slice(n())
worst_young <- worst[["2–4"]]; best_young <- best[["2–4"]]
```

Two things show up that the quartile summary hid.

**The model's ranking works better for older children than younger ones.** The item-level
correlation runs
`r paste(sprintf("%s %+.2f", cr$band, cr$r), collapse = ", ")`. For the 2–4s the prediction
carries little item-level information, as expected given that many are near chance on
every item, which flattens the range over which the correlation is computed.

**The calibration gap flips with age.** Mean observed minus mean predicted is
`r paste(sprintf("%s %+.1f pts", cr$band, 100*cr$gap), collapse = ", ")`. Every band still
falls short of the adult-derived prediction, but the shortfall shrinks steadily with age:
the youngest miss it by `r sprintf('%.0f', -100*cr$gap[1])` points, the oldest by only
`r sprintf('%.0f', -100*cr$gap[3])`. SPoSE was fit to adults, so this is what you would
expect if it describes a mature similarity space that children are still moving toward.

The per-trial extremes are informative. **`r worst$trial_id`**
(*`r worst$odd`*) is predicted at `r pct(worst$predicted)` and the 2–4s get
`r pct(worst_young)`, while **`r best$trial_id`** (*`r best$odd`*) runs the other way,
predicted `r pct(best$predicted)`, observed `r pct(best_young)`. Items like the latter are
the ones to mine when building a bank that works for the youngest children.

# The basic-level tier {#sec-basic}

A second kind of item went live on **21 August 2026**: two photographs of *one* concept
against a different concept from the same category, for example two dogs and a cat. The pair are
different exemplars, so the trial cannot be solved by matching pictures; a child has to see
that two things which look unalike are the same kind. Six open every session and six more are
sampled per child and spread through the test block, taking a session from 51 to 63 trials.

Everything above this section is the vocab-gradient bank only. The two tiers are never pooled.

```{r basic-rollout}
# A partial session on the new design also has fewer than 51 trials, so length
# alone cannot date it. Seeing a basic-level trial can: the tier did not exist
# before the change.
saw_basic <- unique(kdb$participantID)
roll <- kids_all %>%
  filter(n_trials > 0) %>%
  mutate(design = ifelse(participantID %in% saw_basic | n_trials > 51,
                         "new (63 trials)", "old (51 trials)")) %>%
  group_by(design) %>%
  summarise(children = n(), completed = sum(complete),
            completion = sum(complete) / n(), .groups = "drop")
kable(roll, digits = 2,
      caption = "Kiosk children by session design. Completion is the number to watch.")
```

```{r basic-first-data}
n_bt <- nrow(kdb); n_bk <- n_distinct(kdb$participantID)
if (n_bt > 0) {
  kdb %>% group_by(participantID, age) %>%
    summarise(trials = n(), correct = sum(correct), accuracy = mean(correct),
              median_rt = median(rt), .groups = "drop") %>%
    kable(digits = 2, caption = "Every basic-level trial a kiosk child has done so far.")
}
```

```{r basic-adults}
if (nrow(adb) > 0) {
  adb %>% group_by(participantID) %>%
    summarise(trials = n(), accuracy = mean(correct), median_rt = median(rt), .groups="drop") %>%
    arrange(desc(median_rt)) %>%
    kable(digits = 2, caption = paste("Non-kiosk sessions that saw the new tier.",
      "The fast ones are deploy smoke tests, not data."))
}
```

::: {.callout-note}
## Insufficient data for interpretation

`r n_bt` basic-level trials from `r n_bk` kiosk `r ifelse(n_bk == 1, "child", "children")`.
No child has yet finished a 63-trial session. What the data *do* confirm is that the tier is
wired up correctly: scoring is consistent on every trial recorded
(`correct` always equals `chosen == odd`), the opening six land at positions 1–6, and the
sampled six are spread through the block as intended.
:::

The first question to check once sessions accumulate is whether these trials function as the
intended floor. The premise is that a child who cannot make the fine-grained discrimination
can still identify the odd item among two dogs and a cat, so basic-level accuracy should sit
**above** vocab-gradient accuracy, particularly for the 2–4s. If it does not, the tier is
not serving the purpose it was added for.

The second is completion. The session grew by 12 trials against a completion rate that was
already `r pct(sum(kids_all$complete) / sum(kids_all$n_trials > 0), 0)`. The opening six may
act in the opposite direction, since an easier start is a plausible retention benefit. The
pre-change cohort in the table above provides the comparison.

# Conclusions

```{r final-numbers}
b_fin <- tidy(m_age, effects="fixed") %>% filter(term=="scale(age)")
```

1. **Use every session.** Including partial sessions doubles the child sample
   (`r n_kid_complete` → `r n_kid_started`) and takes the age effect from non-significant to
   β = `r sprintf('%+.2f', b_fin$estimate)`, *p* = `r fp(b_fin$p.value)`. Completion is
   age-graded, so completers-only is an older, non-representative sample.

2. **The age effect is not a vocabulary effect.** Target AoA does not predict child accuracy
   (*p* = `r fp(aoa_tab$p.value[aoa_tab$model=="B. AoA alone"])`), does not interact with age
   (*p* = `r fp(a_int$p.value)`), and does not shift the age estimate at all. Whatever is
   developing is the conceptual comparison rather than word knowledge, so the next design
   iteration need not prioritise vocabulary controls.

3. **Screen trials, not children.** A `r RT_FLOOR` ms trial-level floor removes fast taps
   while keeping every child's considered responses. This retains more data than dropping five
   children entirely, and allows deliberation to enter the model as a covariate.

4. **Shorten the session, or reprofile it.** The commonest stopping point is trial 8, exactly
   where warmup ends and test trials begin. Interleaving easy items, or cutting the session
   below `r RUN_LEN` trials, should raise completion where it is lowest, at ages 2–4.

5. **The conflict prediction is still open.** Interaction *p* = `r fp(i_p)` with the full
   sample. Both groups sit off floor and ceiling, so this is answerable with more children
   rather than a redesign.

6. **Restart the adult arm.** Nothing on Prolific since 14 August; the adult baseline is
   `r n_ad` participants.

```{r session-info}
cat("Data pulled: ", format(file.mtime(params$trials), "%Y-%m-%d %H:%M"), "\n")
cat("Children:    ", n_kid_all, "consented,", n_kid_started, "answered >=1 trial,",
    n_kid_complete, "finished\n")
cat("Adults:      ", n_ad, "usable\n")
cat("Trials:      ", nrow(kdf), "child +", nrow(adf), "adult (>=", RT_FLOOR, "ms)\n")
```
