---
title: "MOCHI Kids — pilot sanity check"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    code_folding: hide
    df_print: paged
    self_contained: true
params:
  refresh: false
  qa_filter: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 9, fig.height = 4.5, fig.align = "center",
  dpi = 110
)
library(tidyverse)
library(scales)
library(here)
library(kableExtra)
library(DT)

if (params$refresh) {
  system2("python3", c(here("analysis", "fetch_data.py")))
}

sessions <- read_csv(here("analysis", "data", "sessions.csv"),
                     show_col_types = FALSE)
trials   <- read_csv(here("analysis", "data", "trials.csv"),
                     show_col_types = FALSE)

age_order  <- c("2","3","4","5","6","7","8","9","10+","adult")
tier_order <- c("training","warmup","familiar","novel","catch")
sessions <- sessions %>% mutate(consent_age = factor(consent_age, levels = age_order))
trials   <- trials %>%
  mutate(consent_age = factor(consent_age, levels = age_order),
         tier        = factor(tier, levels = tier_order))

# Helper: consistent kable styling for static tables.
ktab <- function(x, caption = NULL, digits = 3, full_width = FALSE) {
  knitr::kable(x, caption = caption, digits = digits) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"),
      full_width = full_width,
      position = "left",
      font_size = 13
    )
}

n_sessions_total <- nrow(sessions)
n_trials_total   <- nrow(trials)
expected_trials  <- if (file.exists(here("public", "manifest.json"))) {
  length(jsonlite::fromJSON(here("public", "manifest.json"))$trials$trial_id)
} else NA_integer_
```

```{r empty-guard, eval = (n_sessions_total == 0), echo = FALSE, results = "asis"}
cat("**No real sessions in the database yet.** Run a few playthroughs (or wait until pilot data lands) and re-knit.")
knitr::knit_exit()
```

> **Tip.** Re-pull from Mongo by knitting with `params=list(refresh=TRUE)`.
> Show flagged sessions with `params=list(qa_filter=FALSE)`.

# QA flags

Each session is flagged if it looks like a button-masher / position-spammer / disengaged participant. `qa_pass = TRUE` means none of the heuristics fired.

```{r qa-table}
qa_cols <- sessions %>%
  select(participantID, consent_age, n_trials, overall_acc, training_acc,
         catch_acc, max_position_share, max_run_length,
         n_fast_lt_500ms, n_stuck_gt_15s, qa_pass)

datatable(
  qa_cols,
  rownames = FALSE,
  filter = "top",
  options = list(pageLength = 15, dom = "tip", scrollX = TRUE),
  class = "compact stripe hover"
) %>%
  formatPercentage(c("overall_acc","training_acc","catch_acc","max_position_share"), 1) %>%
  formatStyle("qa_pass",
              backgroundColor = styleEqual(c(TRUE, FALSE), c("#d4edda", "#f8d7da")))
```

```{r flagged-count, results="asis"}
n_pass <- sum(sessions$qa_pass)
n_flag <- nrow(sessions) - n_pass
cat(sprintf("**%d/%d sessions passed QA.** %d flagged.\n",
            n_pass, nrow(sessions), n_flag))
```

```{r filter-bad-actors, echo = FALSE}
if (isTRUE(params$qa_filter)) {
  good_pids <- sessions %>% filter(qa_pass) %>% pull(participantID)
  sessions  <- sessions %>% filter(participantID %in% good_pids)
  trials    <- trials   %>% filter(participantID %in% good_pids)
}
n_sessions <- nrow(sessions)
n_trials   <- nrow(trials)
```

```{r filter-status, results="asis"}
if (params$qa_filter) {
  cat(sprintf("Downstream sections show **qa_pass only** (%d kept, %d removed).\n",
              n_sessions, n_sessions_total - n_sessions))
} else {
  cat("QA filter is **off** — all sessions included.\n")
}
```

```{r post-filter-empty, eval = (n_sessions == 0), echo = FALSE, results = "asis"}
cat("\n**No sessions left after QA filtering.** Re-knit with `params=list(qa_filter=FALSE)` to inspect flagged sessions.\n")
knitr::knit_exit()
```

# Sample

`r n_sessions` sessions, `r format(n_trials, big.mark=",")` trials.

```{r sample-by-age}
sessions %>%
  count(consent_age, .drop = FALSE) %>%
  ggplot(aes(consent_age, n)) +
  geom_col(fill = "#6ec1e4") +
  geom_text(aes(label = n), vjust = -0.3, size = 4) +
  labs(x = "self-reported age", y = "sessions",
       title = "Who's played so far") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

# Completion

A complete session has `n_trials == `r expected_trials``. Sessions saved during older manifest versions may show a different total (74, 80, …).

```{r completion}
sessions %>%
  ggplot(aes(n_trials)) +
  geom_histogram(binwidth = 1, fill = "#ff6f61") +
  geom_vline(xintercept = expected_trials, linetype = "dashed", color = "grey30") +
  annotate("text", x = expected_trials, y = Inf, label = paste0("full: ", expected_trials),
           vjust = 1.5, hjust = -0.05, color = "grey30") +
  labs(x = "trials completed", y = "sessions",
       title = "Completion histogram") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))

incomplete <- sessions %>% filter(n_trials < expected_trials)
```

```{r incomplete-table, eval = nrow(incomplete) > 0}
incomplete %>%
  select(participantID, consent_age, n_trials, n_correct) %>%
  ktab(caption = "Sessions that didn't finish")
```

# Accuracy by tier

Adult ceilings on the curated set: training≈1, warmup≈1, familiar~0.97, novel~0.91. Chance for 3-AFC is 1/3.

```{r acc-tier-table}
trials %>%
  group_by(tier) %>%
  summarise(n = n(),
            acc = mean(correct, na.rm = TRUE),
            .groups = "drop") %>%
  ktab(caption = "Trial-level accuracy by tier")
```

```{r acc-tier-plot}
trials %>%
  group_by(participantID, consent_age, tier) %>%
  summarise(acc = mean(correct, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(tier, acc, color = consent_age)) +
  geom_hline(yintercept = 1/3, linetype = "dashed", color = "grey50") +
  geom_jitter(width = 0.18, height = 0, size = 3, alpha = 0.85) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(x = NULL, y = "accuracy",
       title = "Per-participant accuracy by tier",
       caption = "dashed line = chance (1/3)",
       color = "age") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

# Reaction time

Sanity: kids should mostly land between 500ms and ~10s. Floor (button-mash) and ceiling (stuck) suggest UI issues.

```{r rt-violin}
trials %>%
  filter(!is.na(rt)) %>%
  ggplot(aes(tier, rt / 1000, fill = tier)) +
  geom_violin(alpha = 0.55, color = NA) +
  geom_boxplot(width = 0.12, fill = "white", outlier.size = 0.8, alpha = 0.9) +
  scale_y_continuous(trans = "log10",
                     breaks = c(0.1, 0.25, 0.5, 1, 2, 5, 10, 30),
                     labels = c("0.1","0.25","0.5","1","2","5","10","30")) +
  labs(x = NULL, y = "RT (seconds, log scale)",
       title = "RT distribution by tier") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold"))
```

```{r rt-summary}
trials %>%
  group_by(tier) %>%
  summarise(median_rt_s   = median(rt, na.rm = TRUE) / 1000,
            p25_rt_s      = quantile(rt, 0.25, na.rm = TRUE) / 1000,
            p75_rt_s      = quantile(rt, 0.75, na.rm = TRUE) / 1000,
            n_under_500ms = sum(rt < 500, na.rm = TRUE),
            n_over_15s    = sum(rt > 15000, na.rm = TRUE),
            .groups = "drop") %>%
  ktab(caption = "RT summary by tier (seconds)", digits = 2)
```

# Per-trial accuracy

If everyone (or no one) gets a specific trial right, the stimulus may be broken or trivially easy. Compare kid accuracy to the adult baseline from the manifest.

```{r per-trial-scatter, fig.height = 6}
per_trial <- trials %>%
  filter(tier %in% c("warmup","familiar","novel")) %>%
  group_by(trial_id, tier, dataset, human_avg_adult) %>%
  summarise(n = n(), kid_acc = mean(correct, na.rm = TRUE), .groups = "drop")

per_trial %>%
  ggplot(aes(human_avg_adult, kid_acc, color = tier)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey60") +
  geom_hline(yintercept = 1/3, linetype = "dotted", color = "grey60") +
  geom_point(size = 2.6, alpha = 0.8) +
  scale_x_continuous(labels = percent_format(), limits = c(0, 1)) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(x = "adult accuracy (manifest)",
       y = "kid accuracy (this dataset)",
       title = "Per-trial: kids vs adults",
       caption = "dashed = identity, dotted = chance") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

```{r odd-trials, results="asis"}
odd <- per_trial %>%
  filter(n >= 3, kid_acc <= 1/3 - 0.05 | kid_acc >= 0.99) %>%
  arrange(kid_acc)
if (nrow(odd) > 0) {
  cat("### Trials worth eyeballing\n\nKids near chance OR everyone correct (n ≥ 3 kids per trial):\n\n")
  print(ktab(odd, digits = 3))
} else {
  cat("_No trial outliers at the n ≥ 3 threshold yet._\n")
}
```

# Position bias

If kids systematically tap a single position regardless of where the oddity is, something's off (visual layout, audio cuing, etc.).

```{r position-table}
trials %>%
  filter(!is.na(chosen_display_pos)) %>%
  count(chosen_display_pos) %>%
  mutate(prop = n / sum(n)) %>%
  ktab(caption = "Tap distribution across positions", digits = 3)
```

```{r position-plot}
trials %>%
  filter(!is.na(chosen_display_pos)) %>%
  count(chosen_display_pos, correct) %>%
  ggplot(aes(factor(chosen_display_pos), n, fill = correct)) +
  geom_col(position = "fill", color = "white") +
  scale_y_continuous(labels = percent_format()) +
  scale_fill_manual(values = c(`TRUE` = "#3aa53a", `FALSE` = "#d9534f")) +
  labs(x = "display position chosen (0=left, 1=mid, 2=right, 3=far-right for n=4 trials)",
       y = "share of taps",
       fill = "correct?",
       title = "Accuracy by chosen position") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

# Schema sanity

Any rows with missing fields that should be there.

```{r schema-session, results="asis"}
session_missing <- sessions %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "field", values_to = "n_missing") %>%
  filter(n_missing > 0)
if (nrow(session_missing)) {
  cat("### Session-level missing fields\n\n")
  print(ktab(session_missing))
} else {
  cat("_All session-level fields populated._\n")
}
```

```{r schema-trials, results="asis"}
trial_missing <- trials %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "field", values_to = "n_missing") %>%
  filter(n_missing > 0)
if (nrow(trial_missing)) {
  cat("### Trial-level missing fields\n\n")
  print(ktab(trial_missing))
} else {
  cat("_All trial-level fields populated._\n")
}
```
