---
title: "THINGS-Kids — stimulus review"
subtitle: "Every trial, the intended answer, and why"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 2
    theme: flatly
    df_print: paged
    self_contained: false
params:
  # "core" is the 20 trials every child sees; "all" adds the 400 rotating
  # bank trials and makes a very long document.
  scope: "core"
  pilot: "pilot_data/pilot_trials.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.show = "hold", out.width = "31%")
library(dplyr)
library(knitr)
library(ggplot2)

d <- read.csv("stimulus_reasoning.csv", stringsAsFactors = FALSE)

# Pilot responses, if any have been collected. Trials nobody has seen simply
# come through as NA rather than being dropped.
pilot_path <- params$pilot
pilot <- if (file.exists(pilot_path)) read.csv(pilot_path, stringsAsFactors = FALSE) else NULL
if (!is.null(pilot)) {
  seen <- pilot %>% group_by(trial_id) %>%
    summarise(n_seen = n(), n_correct = sum(correct), .groups = "drop") %>%
    mutate(pilot_acc = n_correct / n_seen)
  d <- left_join(d, seen, by = "trial_id")
} else {
  d$n_seen <- NA_integer_; d$pilot_acc <- NA_real_
}

img <- function(concept) file.path("..", "public", "stimuli", paste0(concept, ".jpg"))

# Plain-language account of why the intended answer is the intended answer.
# SPoSE similarity is on 0-1; the softmax P is the model's predicted chance
# that an adult picks the intended item.
reason <- function(r) {
  if (r$duplicate_image == 1) {
    return(sprintf(
      "**Attention check.** Two copies of the same photo (*%s*) plus *%s*. The duplicate is the point — anyone attending cannot miss it.",
      r$pair_a, r$odd_concept))
  }
  agree <- r$clip_margin > 0
  sprintf(paste0(
    "**%s** is the odd one out. SPoSE places *%s* and *%s* at **%.2f** (%s), ",
    "while *%s* (%s) sits at only %.2f and %.2f from them. ",
    "The model puts an adult's chance of picking it at **%.0f%%**. ",
    "CLIP %s: the visual margin is **%+.3f**, so appearance %s."),
    toupper(r$odd_concept), r$pair_a, r$pair_b, r$spose_pair,
    ifelse(nzchar(r$pair_category), r$pair_category, "uncategorised"),
    r$odd_concept, ifelse(nzchar(r$odd_category), r$odd_category, "uncategorised"),
    r$spose_odd_a, r$spose_odd_b, 100 * r$p_correct_spose,
    ifelse(agree, "agrees", "disagrees"), r$clip_margin,
    ifelse(agree, "points at the same answer",
           "pulls *against* the intended answer — this is a conflict trial"))
}

# Emit raw HTML rather than knitr::include_graphics(): inside a
# results="asis" loop include_graphics() returns an object that is never
# printed, so the images silently vanish. Raw <img> also lets the intended
# answer carry a visible outline.
show_trial <- function(r) {
  flag <- ""
  if (!is.na(r$pilot_acc)) {
    flag <- if (r$pilot_acc < 0.5)
      "<span class='miss'>&#10007; missed in pilot</span>"
    else "<span class='hit'>&#10003; pilot correct</span>"
  }
  cat(sprintf("\n\n#### %s %s\n\n", r$trial_id, flag))
  # No position is quoted: the app reshuffles the three images independently
  # for every child, so "position 3" here would mean nothing on screen.
  cat(sprintf("<p class='meta'>%s &middot; %s</p>\n\n",
              r$condition,
              ifelse(nzchar(r$difficulty_bin), r$difficulty_bin, r$tier)))
  cat(reason(r), "\n\n")

  concepts <- c(r$img_1, r$img_2, r$img_3)
  odd_pos <- r$oddity_index + 1
  cat("<div class='triad'>\n")
  for (k in seq_along(concepts)) {
    is_odd <- (k == odd_pos)
    cat(sprintf(
      "<figure class='%s'><img src='%s' alt='%s'><figcaption>%s%s</figcaption></figure>\n",
      if (is_odd) "cell odd" else "cell",
      img(concepts[k]), concepts[k], concepts[k],
      if (is_odd) " &larr; odd one out" else ""))
  }
  cat("</div>\n\n")
}
```

<style>
.triad { display: flex; gap: 10px; margin: 10px 0 26px; flex-wrap: wrap; }
.triad .cell { margin: 0; flex: 1 1 180px; max-width: 240px; }
.triad img { width: 100%; aspect-ratio: 1; object-fit: cover;
             border-radius: 8px; display: block; border: 3px solid transparent; }
.triad .odd img { border-color: #eb6834; }
.triad figcaption { font-size: 12px; color: #666; margin-top: 5px; text-align: center; }
.triad .odd figcaption { color: #c0521f; font-weight: 600; }
p.meta { font-size: 12px; color: #888; margin: -6px 0 8px; text-transform: uppercase;
         letter-spacing: .06em; }
span.miss { color: #c0392b; font-size: 13px; font-weight: 600; }
span.hit  { color: #27ae60; font-size: 13px; font-weight: 500; }
h4 { margin-top: 34px; border-top: 1px solid #eee; padding-top: 16px; }
</style>

## How to read this

Each trial shows three photographs in the order a child sees them. Two are
meant to go together; the third is the intended answer.

**The intended answer comes from SPoSE**, a model fit to 4.7 million adult
odd-one-out judgements (Hebart et al., 2023). It is not a ground truth — it is
a well-supported prediction about what adults *would* say, and part of the job
of this document is to find the trials where that prediction looks wrong.

Two numbers carry most of the reasoning:

- **SPoSE similarity** (0–1) between concepts. The pair should be high; the odd
  item should be low against both.
- **CLIP visual margin.** Positive means the photographs agree with SPoSE.
  Negative means they disagree — the intended odd item actually *looks* like it
  belongs, and only category knowledge supports the answer. Half of all test
  trials are deliberately negative.

```{r overview}
d %>%
  filter(section %in% c("core", "block")) %>%
  count(condition, difficulty_bin) %>%
  tidyr::pivot_wider(names_from = condition, values_from = n, values_fill = 0) %>%
  kable(caption = "Test trials by condition and difficulty (core + full bank)")
```

## Pilot: where the model and the adult disagreed

```{r pilot-summary, eval = !is.null(pilot)}
pd <- d %>% filter(!is.na(pilot_acc), section %in% c("core", "block"))
pd %>%
  group_by(condition) %>%
  summarise(trials = n(), correct = sum(n_correct), acc = mean(pilot_acc), .groups = "drop") %>%
  kable(digits = 2, caption = "Adult pilot accuracy by condition")
```

```{r pilot-calibration, eval = !is.null(pilot), fig.width = 7, fig.height = 4.2, out.width = "88%"}
ggplot(pd, aes(p_correct_spose, pilot_acc)) +
  geom_hline(yintercept = 1/3, linetype = "dashed", colour = "grey60") +
  geom_abline(slope = 1, intercept = 0, colour = "grey75") +
  geom_jitter(aes(colour = condition), height = 0.03, width = 0, size = 2.6, alpha = 0.85) +
  scale_colour_manual(values = c(visual_agree = "#2a78d6", visual_conflict = "#eb6834")) +
  scale_x_continuous(limits = c(0.3, 1)) + scale_y_continuous(limits = c(-0.05, 1.05)) +
  labs(x = "SPoSE predicted P(correct)", y = "Adult pilot accuracy",
       title = "Calibration: does the model predict what the adult actually did?",
       subtitle = "Grey line is perfect calibration; dashed line is chance") +
  theme_minimal(base_size = 12) + theme(legend.position = "bottom")
```

Points far **below** the grey line are the ones to worry about: trials the model
calls easy that a fully attentive adult still got wrong. Those are candidates for
removal, because a trial no adult endorses cannot measure anything in a child.

```{r pilot-misses, eval = !is.null(pilot)}
pd %>%
  filter(pilot_acc < 0.5, p_correct_spose > 0.75) %>%
  arrange(desc(p_correct_spose)) %>%
  select(trial_id, condition, pair_a, pair_b, odd_concept,
         spose_pair, p_correct_spose, clip_margin) %>%
  kable(digits = 2,
        caption = "Model says easy, adult said otherwise — review these first")
```

## Core trials

Every child sees these 20, so they carry the age comparison and deserve the
closest reading.

```{r core, results = "asis"}
core <- d %>% filter(section == "core") %>% arrange(desc(p_correct_spose))
invisible(lapply(seq_len(nrow(core)), function(i) show_trial(core[i, ])))
```

## Training, warmup, and attention checks

```{r intro, results = "asis"}
intro <- d %>% filter(section %in% c("training", "warmup", "catch")) %>%
  arrange(factor(section, levels = c("training", "warmup", "catch")), trial_id)
invisible(lapply(seq_len(nrow(intro)), function(i) show_trial(intro[i, ])))
```

## Rotating bank

```{r bank, results = "asis", eval = params$scope == "all"}
bank <- d %>% filter(section == "block") %>% arrange(block, trial_id)
for (b in unique(bank$block)) {
  cat(sprintf("\n\n### Block %s\n", b))
  bb <- bank %>% filter(block == b)
  invisible(lapply(seq_len(nrow(bb)), function(i) show_trial(bb[i, ])))
}
```

```{r bank-note, results = "asis", eval = params$scope != "all"}
cat(sprintf(paste0(
  "The 400 rotating bank trials are omitted here to keep the document readable. ",
  "Each child sees one 20-trial block of the %d currently active.\n\n",
  "Render the full set with:\n\n",
  "```r\nrmarkdown::render(\"stimulus_review.Rmd\", params = list(scope = \"all\"))\n```\n"),
  length(unique(d$block[d$section == "block" & d$block != ""]))))
```

## Provenance

All photographs come from the vocab-gradient published item set (THINGS,
copyright-free release), restricted to concepts appearing in `test_items.csv`
so results join to that paper's 4AFC item data. Similarity terms are
recomputed by `build_reasoning_table.py` from the same matrices the bank was
built with: SPoSE 66-d (Hebart et al., 2023) and CLIP ViT-B/32 with OpenAI
weights, matching the model the vocab-gradient paper used.

No image appears twice within a session, and no two concepts in different
trials of one session exceed CLIP .88 — near-identical photographs of
different concepts read as repeats even when nothing is literally repeated.
